---
title: "BayesSimwithShortvLongtwithNumericalmethods"
author: "Naveed Merchant"
date: "November 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Long Tail Experiments

It has been noticed that the Gaussian Kernel is unstable if the data comes from a very long tail distribution.

If the data comes from a Cauchy distribution, the bayes factor, in general, is unstable.

We verify that switching Kernels, in general, seems to fix problems.

We try to show this in plots and by using "BayesSim"

We now show the "Cauchy v Gaussian problem"

```{r setup and loading}
library(parallel)
library(ggplot2)
source("MarginalLikIntfunctions.R")
source("Laplacefunction.R")
set.seed(1000)

p = rbeta(500,.5,.5)
dlength = 400

dataset1 <- rcauchy(dlength)
dataset2 <- matrix(data = NA, nrow = length(p),ncol = dlength)
for(i in 1:length(p))
{
  unifdraw = runif(dlength)
  for(j in 1:dlength)
  {
    if(unifdraw[j] > p[i])
    {
      dataset2[i,j] = rcauchy(1)
    }
    else
    {
      dataset2[i,j] = rnorm(1, mean = 0, sd = 1.4826)
    }
  }
}

XT1 <- dataset1[1:(length(dataset1)*.3)]
XV1 <- dataset1[-(1:(length(dataset1)*.3))]

likvec = function(h) {sum(log(HallKernel(h,datagen2 = XT1, x = XV1)))}
bwlik = optimize(f = function(h){  likvec(h)}, lower = 0, upper = 10, maximum = TRUE)
ExpectedKernML1 = laplace.kernH2(y = XT1, x = XV1, hhat = bwlik$maximum)

CL1 <- makeCluster(8)
clusterExport(cl = CL1, list("XT1","XV1","dataset2","logmarg.specialkernMCimport", "ExpectedKernML1", "dlength", "logSumExp", "HallKernel", "laplace.kernH", "logintegrand.Hall", "loglike.KHall", "KHall", "laplace.kernH2", "hessian"), envir = environment())
result <- parApply(cl = CL1, dataset2[1:500,], 1, FUN = function(x){
  XT2 = x[1:(dlength*.3)]
  XV2 = x[-(1:(dlength*.3))] 
  likvec = function(h) {sum(log(HallKernel(h,datagen2 = XT2, x = XV2)))}
  bwlik = optimize(f = function(h){  likvec(h)}, lower = 0, upper = 10, maximum = TRUE)
  ExpectedKernML2 = laplace.kernH2(y = XT2, x = XV2, hhat = bwlik$maximum)
  
  likvec2 = function(h) {sum(log(HallKernel(h,datagen2 = c(XT1,XT2), x = c(XV1,XV2))))}
  bwlik2 = optimize(f = function(h){  likvec2(h)}, lower = 0, upper = 10, maximum = TRUE)
  
  ExpectedKernMLcomb = laplace.kernH2(y = c(XT1,XT2), x = c(XV1,XV2), hhat = bwlik2$maximum)
  
  return(ExpectedKernML1[1] + ExpectedKernML2[1] - ExpectedKernMLcomb[1])
})
stopCluster(CL1)
# lm4 <- lm(result~p)
# lw1 = loess(result~p)
# j <- order(p)
# plot(p,result, xlab = "p", ylab = "Log Bayes Factor", pch = '*')
# lines(p[j],lw1$fitted[j],col="red",lwd=3)
# 
# 
# abline(a = 0, b = 0)

dfCVBF = data.frame(logBF = result, p = p)
ggplot(dfCVBF, aes(x = p, y = logBF)) + geom_point()  + geom_hline(yintercept = 0, color = "blue") + geom_smooth(colour = "red")


#abline(lm4, col = "blue")

sorteddfCVBF = dfCVBF[order(dfCVBF$p),]
nonparvarest = 0
for(j in 1:(nrow(sorteddfCVBF) - 1) )
{
  nonparvarest = nonparvarest + (sorteddfCVBF[j + 1, 1] - sorteddfCVBF[j, 1])^2
}
nonparvarest = nonparvarest / (2* nrow(sorteddfCVBF))

#A variance estimate

nonparvarest

#A standard deviation estimate

sqrt(nonparvarest)


```

We are examining the log BF, of the test that checks whether a standard Cauchy is the same as a mixture of a standard Cauchy and a normal whose IQR is roughly the same as the Cauchy.


```{r KStest equivalent}
kslist = c()
for(j in 1:500)
{
  kslist[j] =  ks.test(dataset1,dataset2[j,])$p
}
# plot(p[1:500],log(kslist), xlab = "p", ylab = "log p-values from KS test", pch = '*')
# abline(a=log(0.05),b=0, col = "blue")

dfKS = data.frame(logks = log(kslist), p = p)
ggplot(dfKS, aes(x = p, y = logks)) + geom_point()  + geom_hline(yintercept = log(.05), color = "blue")



```

We also compare this to the BFs produced by the Polya tree test.

We center and scale before proceeding. The below file code deals with the normal distribution being used as a base distribution.


```{r comparison to polya tree test}
PTlist = c()
for(j in 1:500)
{
  mediancombineddataset = median(c(dataset1, dataset2[j,]))
  IQRcombineddataset = IQR(c(dataset1, dataset2[j,]))
  dataset1adj = (dataset1 - mediancombineddataset) / (IQRcombineddataset / 1.36)
  dataset2adj = (dataset2[j,] - mediancombineddataset) / (IQRcombineddataset / 1.36)
  PTlist[j] = PolyaTreetest(dataset1adj,dataset2adj, Ginv = qnorm, c = 1, leveltot = 10)$logBF
}
# plot(p[1:500],PTlist, xlab = "p", ylab = "log BF values from Polya Tree test", pch = '*')
# abline(a=0,b=0, col = "blue")


dfPT = data.frame(logBF = PTlist, p = p)
ggplot(dfPT, aes(x = p, y = logBF)) + geom_point()  + geom_hline(yintercept = 0, color = "blue") + geom_smooth(colour = "red")

sorteddfPT = dfPT[order(dfPT$p),]
nonparvarest = 0
for(j in 1:(nrow(sorteddfPT) - 1) )
{
  nonparvarest = nonparvarest + (sorteddfPT[j + 1, 1] - sorteddfPT[j, 1])^2
}
nonparvarest = nonparvarest / (2* nrow(sorteddfPT))

#Variance estimate

nonparvarest

#Standard deviation estimate

sqrt(nonparvarest)

```

How does this compare to preforming the same test but using a Cauchy quantile instead?

```{r comparison to polya tree test2}
PTlist2 = rep(0, times = 500)
for(j in 1:500)
{
  mediancombineddataset = median(c(dataset1, dataset2[j,]))
  IQRcombineddataset = IQR(c(dataset1, dataset2[j,]))
  dataset1adj = (dataset1 - mediancombineddataset) / (IQRcombineddataset / 1.36)
  dataset2adj = (dataset2[j,] - mediancombineddataset) / (IQRcombineddataset / 1.36)
  PTlist2[j] = PolyaTreetest(dataset1adj,dataset2adj, Ginv = qcauchy, c = 1, leveltot = 10)$logBF
}
# plot(p[1:500],PTlist2, xlab = "p", ylab = "log BF values from Polya Tree test", pch = '*')
# abline(a=0,b=0, col = "blue")
# abline(a=-2.99,b=0, col = "green")

dfPT2 = data.frame(logBF = PTlist2, p = p)
ggplot(dfPT2, aes(x = p, y = logBF)) + geom_point()  + geom_hline(yintercept = 0, color = "blue") + geom_smooth(colour = "red")

sorteddfPT = dfPT2[order(dfPT$p),]
nonparvarest = 0
for(j in 1:(nrow(sorteddfPT) - 1) )
{
  nonparvarest = nonparvarest + (sorteddfPT[j + 1, 1] - sorteddfPT[j, 1])^2
}
nonparvarest = nonparvarest / (2* nrow(sorteddfPT))

#Variance estimate

nonparvarest

#Standard deviation estimate

sqrt(nonparvarest)


```

We can now compare all plots. The Polya tree that uses a Cauchy base distribution is variable but ascends above 0 corecctly. The Polya tree that uses a normal base distribution is also variable but crosses 0 late.




```


